The speed of PyPy

### 2010-02-17 talk/pycon2010/pypyspeed/talk.tex
work on slides. They need a bit more work, but sort-of-ok-now
(hpk, fijal) minor stuff
(hpk, fijal) progress

\begin{frame}
  \frametitle{JIT - what's that about?}
  \begin{itemize}
    \item {\bf JIT is not a magical device!}
      \pause
    \item removes bytecode overhead
    \item removes frame overhead
    \item can make runtime decisions
    \item more classic optimization that can follow
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{How well it works in practice?}
\end{frame}

@@

\begin{frame}[fragile]
  \frametitle{Removing frame overhead}
  \begin{verbatim}
x = y + z
  \end{verbatim}
  \begin{itemize}
    \item above has 5 frame accesses
    \item they can all be removed (faster!)
      \pause
    \item they prevent optimizations from happening
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Removing object boxing}
  \begin{verbatim}
i = 0
while i < 100:
  i += 1
  \end{verbatim}
  \begin{itemize}
    \item for each iteration we do a comparison and addition
    \item xxx integers on valuestack and xxx integers in locals
    \item all of those can be removed
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Access costs}
  \begin{itemize}
    \item local access costs nothing
    \item global access is cheap, if you don't change global {\ttfamily \_\_dict\_\_} too much XXX rephrase
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Frame escapes}
  \begin{itemize}
    \item calling {\ttfamily sys.\_getframe()}, {\ttfamily sys.exc\_info()}
    \item exception escaping
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Shared dicts (aka hidden classes)}
  \begin{itemize}
    \item instance {\ttfamily \_\_dict\_\_ } lookup becomes a list lookup
    \item if you're evil, it'll bail back to dict lookup
      \pause
    \item only for newstyle classes!
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Version tags}
  \begin{itemize}
    \item dicts on types are version-controlled
    \item this means methods are usually in known places
      \pause
    \item ... if you don't modify them too often
    \item counters on classes are bad
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Call costs}
  \begin{itemize}
    \item calls can be inlined
    \item simple arguments are by far the best
    \item avoid {\ttfamily *args} and {\ttfamily **kwds}
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Future}
  \begin{itemize}
    \item release end March
    \item to try it out
  \end{itemize}
\end{frame}

>>>>

\begin{frame}
  \frametitle{How fast is PyPy?}
  \pause
  \begin{itemize}
    \item pretty fast, in places
    \item slower than cpython in other places
    \item overall, it depends
    \item graphs
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{JIT - what's that about?}
  \begin{itemize}
      \pause
    \item {\bf JIT is not a magical device!}
      \pause
    \item removes bytecode overhead
    \item removes frame overhead
    \item can make runtime decisions
    \item more classic optimization that can follow
  \end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
  \frametitle{Removing frame overhead}
  \begin{verbatim}
x = y + z
  \end{verbatim}
  \begin{itemize}
    \item above has 5 frame accesses
    \item they can all be removed (faster!)
      \pause
    \item {\bf this enables further optimizations}
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Removing object boxing}
  \begin{verbatim}
i = 0
while i < 100:
  i += 1
  \end{verbatim}
  \begin{itemize}
    \item for each iteration we do a comparison and addition
    \item xxx integers on valuestack and xxx integers in locals
    \item all boxing can be removed
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Variable access costs}
  \begin{itemize}
    \item local access costs nothing
    \item global access is cheap, if you don't change global {\ttfamily \_\_dict\_\_} too much XXX rephrase
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Frame escapes}
  \begin{itemize}
    \item {\bf JIT} normally removes frame overhead, but
    \item calling {\ttfamily sys.\_getframe()}, {\ttfamily sys.exc\_info()}
    \item exception escaping
    \item prevents a lot of optimizations
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Shared dicts (aka hidden classes)}
  \begin{itemize}
    \item instance {\ttfamily \_\_dict\_\_ } lookup becomes an array lookup
    \item if you're evil, it'll bail back to dict lookup
      \pause
    \item {\bf only for newstyle classes}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Version tags}
  \begin{itemize}
    \item dicts on types are version-controlled
    \item this means method lookup can be removed
      \pause
    \item ... if you don't modify them too often
    \item counters on classes are bad
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Call costs}
  \begin{itemize}
    \item calls can be inlined
    \item simple arguments are by far the best
    \item avoid {\ttfamily *args} and {\ttfamily **kwds}
    \item however, {\ttfamily f(a=3, b=c)} is fine
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Future}
  \begin{block}{release end March}
    \begin{itemize}
      \item will contain a working JIT
      \item will not speed up all cases
      \item might eat all your memory
    \end{itemize}
  \end{block}
\end{frame}
